\textsf{Are large language models superhuman chemists?}

### 2024-10-24 src/tex/ms.tex
Update src/tex/ms.tex

While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions). In addition, \chembench samples different skills on various difficulty level: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations). We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}

\paragraph{\enquote{Tiny} subset}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
To address this, we also provide a subset (\variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus. While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
Our human volunteers answered all the questions in this subset.

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on the \enquote{tiny} subset of the  \chembench corpus.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions).
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

\paragraph{Volunteers}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those,
\variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} have a bachelor's degree. For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.

\paragraph{Comparison with models}
For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans. The performance metrics reported for models in the main text are computed on the same questions the humans answered. Metrics for the entire corpus are reported in the appendix.

\subsection{Data annotation}\label{sec:meth-topic}
In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question. We used following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g. the H/P statements of chemicals). Reasoning is required if answering a question requires multiple reasoning steps.
Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.

\section*{Data and code availability}
The code and data for \chembench is available at \url{https://github.com/lamalab-org/chem-bench}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\showyourwork} framework.\autocite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}.
To facilitate reproduction, some intermediate analysis results are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

\section*{Acknowledgements}
This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
A.A.\ gratefully acknowledges financial support for this (publication/project/research) by the Fulbright U.S. Student Program, which is sponsored by the U.S. Department of State and German-American Fulbright Commission. Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.

>>>>

While many existing benchmarks are designed around \gls{mcq}, this does not reflect the reality of chemistry education and research.
For this reason, \chembench samples both \gls{mcq} and open-ended questions (\variable{output/mcq_questions.txt} \gls{mcq} questions and \variable{output/non_mcq_questions.txt} open-ended questions). In addition, \chembench samples different skills on various difficulty levels: From basic knowledge questions (as knowledge underpins reasoning processes\autocite{hu2024towards,bloom1956taxonomy}) to complex reasoning tasks (such as finding out which ions are in a sample given a description of observations). We also include questions about chemical intuition, as showing human-aligned preferences is relevant for applications such as hypothesis generation or optimization tasks.\autocite{zhang2024omniopenendednessmodelshuman}

\paragraph{\chembenchmini}
It is important to note that a smaller subset of the corpus might be more practical for routine evaluations.\autocite{polo2024tinybenchmarks}
For instance,~\textcite{liang2023holistic} report costs of more than \$10,000 for \gls{api} calls for a single evaluation on the widely used \gls{helm} benchmark.
To address this, we also provide a subset (\chembenchmini, \variable{output/num_human_answered_questions.txt} questions) of the corpus that was curated to be a diverse and representative subset of the full corpus. While it is impossible to comprehensively represent the full corpus in a subset, we aimed to include a maximally diverse set of questions and a more balanced distribution of topics and skills (see \Cref{sec:subset-selection} for details on the curation process).
Our human volunteers answered all the questions in this subset.

@@

\begin{figure}[!h]
    \centering
    \includegraphics{figures/human_subset_performance.pdf}
    \caption{\textbf{Performance of models and humans on \chembenchmini.} The figure shows the percentage of questions that the models answered correctly. We use horizontal bars to indicate the performance of various models and highlight statistics of the human performance.
    The evaluation we use here is very strict as it only considers a question answered correctly or incorrectly, partially correct answers are also considered incorrect.
    \Cref{fig:barplot_all_correct_all_questions} provides an overview of the performance of various models on the entire corpus.
    PaperQA2\autocite{skarlinski2024language} is an agentic system that can also search the literature to obtain an answer. We find that the best models outperform all humans in our study when averaged over all questions (even though humans had access to tools such as web search and ChemDraw for a subset of the questions).
    }
    \label{fig:human_vs_models_bar}
    \script{plot_overview_performance_plot.py}
\end{figure}

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close to the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about the \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

@@

\paragraph{Human volunteers}
Users were open to reporting about their experience in chemistry.
Overall, \variable{output/num_users_with_education_info.txt} did so.
Out of those, \variable{output/num_human_postdoc.txt} are beyond a first postdoc, \variable{output/num_human_master.txt} have a master's degree (and are currently enrolled in Ph.D.\ studies), and \variable{output/num_human_bachelor.txt} has a bachelor's degree. For the analysis, we excluded volunteers with less than two years of experience in chemistry after their first university-level course in chemistry.

\paragraph{Comparison with models}
For the analysis, we treated each human as a model. We computed the topic aggregated averages per human for analyses grouped by topic and then averaged over all humans. The performance metrics reported for models in the main text are computed on the same questions the humans answered. Metrics for the entire corpus are reported in the appendix (\Cref{sec:model_performance_app}).

\subsection{Data annotation}\label{sec:meth-topic}
In the curation of our dataset we manually assigned difficulty levels as well as required skills to each question. We used the following guidelines for these annotations: calculation is required if answering a question would require the use of a calculator, knowledge is required if answering a question requires non-trivial knowledge of facts (e.g. the H/P statements of chemicals). Reasoning is required if answering a question requires multiple reasoning steps.
Basic questions are those that only requires those skills up to high-school level. Advanced questions would require an expert multiple minutes up to hours to answer.

\section*{Data and code availability}
The code and data for \chembench are available at \url{https://github.com/lamalab-org/chem-bench}.
The code for the app for our human baseline study is available at \url{https://github.com/lamalab-org/chem-bench-app}.
To ensure reproducibility, this manuscript was generated using the \href{https://show-your.work/en/latest/}{\showyourwork} framework.\autocite{Luger2021}
The code to rebuild the paper (including code for all figures and numbers next to which there is a GitHub icon) can be found at \url{\GitHubURL}.
To facilitate reproduction, some intermediate analysis results are cached at \url{http://dx.doi.org/10.5072/zenodo.34706}.

\section*{Acknowledgements}
This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S. Student Program, which is sponsored by the U.S.\ Department of State and German-American Fulbright Commission. Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.
